Optimal (Multiway) Spatial Joins

### 2024-06-03 msj.tex
yufei

Let $S$ be a set of segments where either all segments are horizontal or all are vertical. Given a rectangle $r$, we define 
\myeqn{
    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \};
    \label{eqn:cross}
}
namely, $\cross_S(r)$ is the set of segments in $S$ crossing $r$.
Let $R$ be a set of rectangles. Given a horizontal segment, we define 
\myeqn{
    \contained_R(h) = \{r \in R \mid \text{$r$ contains $h$} \};
    \label{eqn:contained}
}
namely, $\contained_R(h)$ is the set of rectangles in $R$ containing $h$.

@@

The computation of $\J^\catt_{i,j}$ is an instance of the H-V $k$-SJ problem. Specifically, collect the top-edges of all rectangles of $R_i$ into a set $H$, and collect the left-edges of all the rectangles of $R_j$ into a set $V$. This yields an H-V $k$-SJ instance whose input comprises all the $R_z$ with $z \in [k] \setminus \set{i,j}$,  $H$, and $V$. Each result tuple consists of a rectangle $r_z \in R_z$, for $z \in [k] \setminus \set{i,j}$, a horizontal segment $h \in H$, and a vertical segment $v \in V$ such that $h \cap v \cap \bigcap_{z \in [k] \setminus \set{i,j}} r_z \ne \emptyset$. There is one-one correspondence between the output of the H-V $k$-SJ and $\J^\catt_{i,j}$. Thus, by Lemma~\ref{lmm:hv}, the H-V $k$-SJ can be solved in $O(k) \cdot (F_{k-1}(n, |\J^\catt_{i,j}|) + n \log n + k \cdot |\J^\catt_{i,j}|)$ time. Converting the output into $\J^\catt_{i,j}$ takes another $O(|\J^\catt_{i,j}|)$ time. Applying $|\J^\catt_{i,j}| \le \out$, we know that $\J^\catt_{i,j}$ can be produced in $O(k) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$ time

@@

Fix an integer $k \ge 3$, as well as $k$ integers $n_1, n_2, ..., n_k$ at least 1. Define $n = \sum_{i=1}^k n_i$. An {\em $(n_1, n_2, ..., n_k)$-input} to the $k$-SJ problem comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$ in $\real^2$ such that $|R_i| = n_i$ for each $i \in [k]$. We assume that each rectangle in $R_i$ ($i \in [k]$) is given a distinct ID from $[n_i]$.

@@

Given an $(n_1, n_2, ..., n_k)$-input $\set{R_1, R_2, ..., R_k}$, we execute the algorithm by traversing a root-to-leaf path in $\T$. In general, at an internal node with the instruction $(i, \lambda, \side; i', \lambda', \side')$, we first identify
\myitems{
    \item $r$ as the rectangle with ID $\lambda$ in $R_i$;
    \item $r'$ as the rectangle with ID $\lambda'$ in $R_{i'}$.
}
Then, we compare $\side(r)$ to $\side'(r')$ (e.g., if $\side = \xright$ and $\side = \ybot$, then the comparison is between $\xright(r)$ to $\ybot(r')$). We branch to
\myitems{
    \item the left child if $\side(r) < \side'(r')$;
    \item the middle child if $\side(r) = \side'(r')$;
    \item the right child if $\side(r) > \side'(r')$.
}
At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in that result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$. The algorithm must produce the correct output on all $(n_1, n_2, ..., n_k)$-inputs. Its {\em worst-case cost} is the height of $\T$.

>>>>

Let $S$ be a set of segments where either all segments are horizontal or all are vertical. Given a rectangle $r$, we define 
\myeqn{
    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \};
    \label{eqn:cross}
}
namely, $\cross_S(r)$ is the set of segments in $S$ crossing $r$.
Let $R$ be a set of rectangles. Given a horizontal segment $h$, we define 
\myeqn{
    \contained_R(h) = \{r \in R \mid \text{$r$ contains $h$} \};
    \label{eqn:contained}
}
namely, $\contained_R(h)$ is the set of rectangles in $R$ containing $h$.

@@

The computation of $\J^\catt_{i,j}$ is an instance of the H-V $k$-SJ problem. Specifically, collect the top-edges of all rectangles of $R_i$ into a set $H$, and collect the left-edges of all the rectangles of $R_j$ into a set $V$. This yields an H-V $k$-SJ instance whose input comprises all the $R_z$ with $z \in [k] \setminus \set{i,j}$,  $H$, and $V$. Each result tuple consists of a rectangle $r_z \in R_z$, for $z \in [k] \setminus \set{i,j}$, a horizontal segment $h \in H$, and a vertical segment $v \in V$ such that $h \cap v \cap \bigcap_{z \in [k] \setminus \set{i,j}} r_z \ne \emptyset$. There is one-one correspondence between the output of the H-V $k$-SJ and $\J^\catt_{i,j}$. Thus, by Lemma~\ref{lmm:hv}, the H-V $k$-SJ can be solved in $O(k) \cdot (F_{k-1}(n, |\J^\catt_{i,j}|) + n \log n + k \cdot |\J^\catt_{i,j}|)$ time. Converting the output into $\J^\catt_{i,j}$ takes another $O(k \cdot |\J^\catt_{i,j}|)$ time. Applying $|\J^\catt_{i,j}| \le \out$, we know that $\J^\catt_{i,j}$ can be produced in $O(k) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$ time

@@

Fix an integer $k \ge 3$ and $k$ integers $n_1, n_2, ..., n_k$ at least 1. Define $n = \sum_{i=1}^k n_i$. An {\em $(n_1, n_2, ..., n_k)$-input} to the $k$-SJ problem comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$ in $\real^2$ such that $|R_i| = n_i$ for each $i \in [k]$. We assume that each rectangle in $R_i$ ($i \in [k]$) is given a distinct ID from $[n_i]$.

@@

Given an $(n_1, n_2, ..., n_k)$-input $\set{R_1, R_2, ..., R_k}$, we execute the algorithm by traversing a root-to-leaf path in $\T$. In general, at an internal node with the instruction $(i, \lambda, \side; i', \lambda', \side')$, we first identify
\myitems{
    \item $r$ as the rectangle with ID $\lambda$ in $R_i$;
    \item $r'$ as the rectangle with ID $\lambda'$ in $R_{i'}$.
}
Then, we compare $\side(r)$ to $\side'(r')$ (e.g., if $\side = \xright$ and $\side' = \ybot$, then the comparison is between $\xright(r)$ to $\ybot(r')$). We branch to
\myitems{
    \item the left child if $\side(r) < \side'(r')$;
    \item the middle child if $\side(r) = \side'(r')$;
    \item the right child if $\side(r) > \side'(r')$.
}
At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in that result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$. The algorithm must produce the correct output on all $(n_1, n_2, ..., n_k)$-inputs. Its {\em worst-case cost} is the height of $\T$.
